---
title: "ASSINGMENT1"
output:
  word_document: default
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
---
output:
  word_document: default
  html_document: default
---

## KUNAL SHARMA

#Load dataset packages
```{r}
```
#import the excel file into the global environment
```{r}

data <- read.csv('A:/DATA_SETS/a1-cereals.csv')
```
#head and  tail function 
```{r}
head(data)

tail(data)
```
#Working with functions for Stats
```{r}
min(data$Fat)
max(data$Fiber)
range(data$Vitamins)
mean(data$Weight)
median(data$Protein)
```

# function will give the summary of  data 
```{r}
summary(data)

```
#str function will structure the data 
```{r}
str(data)
```

```{r}
plot(data)
```
#X-Y plot for two quantitative variables(Scatterplot)
```{r}
plot(x= data$Carbohydrates, y=data$Sugars,
     xlab = "Carbs",
     ylab = "SugarContent",
     xlim = c(10,20),
     ylim = c(2,10),
     main = "Carbs with Sugars")

```
#Add some Options
```{r}
plot(data$Carbohydrates,data$Sugars)
```
#Plotting a quantitative variable Audience_Size
```{r}
plot(data$Weight)
```
#LineplotHistogram,Boxplot
```{r}
hist(data$Protein,main = "HISTOGRAM")
boxplot(data$Potassium,data$Vitamins,main="BOXPLOT")
plot(x= data$Carbohydrates, y=data$Sugars,

     xlab = "Carbs",
     ylab = "SugarContent",
	 type="l",main="Example of Sugar and Carbs COntent")

```

#The End




